---
title: "Group 27 CA pollution burden"
author: "Mia Guarnieri, Sarah Rohlfing"
date: "11/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, warning = FALSE, message = FALSE)

#attaching packages
library(tidyverse)
library(here)
library(janitor)
```

### Overview

This data visualization examines pollution levels as related to the percentage of a California county's population that is white. Data was obtained from the CalEnviroScreen 3.0 (or CES 3.0), which is the most recent version of the California Communities Environmental Health Screening Tool - a cumulative impact screening methodology created by the Office of Environmental Health Hazard Assessment at the California Environmental Protection Agency. Per the California Environmental Justice Alliance guidelines, communities with high total CES 3.0 scores have high cumulative pollution burdens and highly vulnerable population characteristics that cause pollutants to become more deadly.

```{r}
# Reading in the data and cleaning up column headers
ces <- read.csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```

### Creating a hexplot to visualize pollution vs. white population in California counties

This basic hexplot (density plot) visualizes CES 3.0 scores as they related to the percentage of the county population that is white. Color corresponds to the density of observations.

```{r}
ggplot(data = ces, aes(x = white, y= ces_3_0_score))+
  geom_hex() +
  scale_fill_gradient(low = "darkseagreen1", high = "grey30")+
  theme_minimal()+
  labs(x = "Percentage of population that is white",
       y = "CES 3.0 score",
       caption = str_wrap("Gradient of CES 3.0 pollution scores by white population percentage. Data: Office of Environmental Health Hazard Assessment, California Environmental Protection Agency"))
```

### Takeaway

This graph shows a negative trend which suggests that communities with a higher percentage of white citizens have lower CES 3.0 scores. This means that higher, more harmful pollution levels occur in more vulnerable communities with larger minority populations. 

### Citation

California Office of Environmental Health Hazard Assessment. 2018. “CalEnviroScreen 3.0.” Oehha.ca.gov, 25 June 2018, https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30. 
